---
title: "Tidying Session"
author: "James"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```

### Read in inverts.xlsx

```{r}
inverts <- read_excel(here("data", 
                           "inverts.xlsx"))

```

### `pivot_longer()` <- tick marks
Use single color (:) to indicate a series of sequential columns A:D
```{r}
inverts_long <- pivot_longer(data = inverts,
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")

# What if you want some in series and some NOT in series?
# c('2016':'2018', '2020')
```

What if we want to get back into wide format?

```{r}
invert_wide <- inverts_long %>%
  pivot_wider(names_from = common_name, 
              values_from = sp_count)
```

### `janitor::clean_names()`

```{r}
inverts_wide <- invert_wide %>% 
  clean_names()
```
### `tidyr::unite()` to combine information from columns

Example: merge the side code and year into a single column called "site_year"

```{r}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year",
        c(site,year),
        sep = "_")
```


### Activity:

Create a new object called 'invert_moyr' starting from inverts_long that unites the month an year columns into a SINGLE column named "mo_yr", using a slash "/" as the separator... then try some other fun separators.


```{r}
inverts_moyr <- inverts_long %>% 
  unite(col = "mo_yr",
        c(month,year),
        sep = "/")
```


### `tidyr::separate()`
```{r}
inverts_sep <- inverts_unite %>% 
  separate(col = site_year, into = c("my_site", "my_year"))
```

### `stringr`+ `mutate()`

Example: using the built in `trees` dataset with mutate() to add 5 to all of the tree heights.

```{r}
trees_new <- trees %>% 
  mutate(height_update = Height + 5)
```

### `str_replace()`

```{r}
ca_abbr <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "california",
                       replacement = "CA"))
```

Example: In the common name column, you want to replace the word "spiny" with the word "awesome". 

`awesome_lobsters`
```{r}
awesome_lobsters <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "spiny",
                       replacement = "awesome"))
```

